#This Script Trains and tests the Random Forest classifiers and saves the results in RDA's. The script was executed using the CHCAA server provided by Interacting Minds Center at Aarhus University.


---
title: "DK_BC"
output: html_document
---
###Creating classifier for backchanneling in Danish
```{r}
#loading packages
install.packages("pacman")
library(pacman)
p_load(gtools, e1071, randomForest, RCurl, caret, tm, quanteda, tidyverse)
#loading data
dk_bc <- read.csv("df_DK.csv")
dk_bc$binary <- dk_bc$Backchannel
```

Cleaning text
```{r}
#Removing punctuation and numbers, as Random Forest can't deal with them
dk_bc$Transcription <-  gsub('[[:punct:] ]+',' ',dk_bc$Transcription)
dk_bc$Transcription <-  gsub('\\d+','tal',dk_bc$Transcription)
```

Assigning wordlists
```{r}
#Column names for result df
coln <- c("Accuracy","Kappa","AccuracyLower","AccuracyUpper","AccuracyNull","AccuracyPValue","McnemarPValue","Sensitivity","Specificity","Pos Pred Value","Neg Pred Value","Precision","Recall","F1","Prevalence","Detection Rate",
          "Detection Prevalence","Balanced Accuracy")
#words and characters the randomForest function cant handle
bad_rf <- c(")", "(", ",", "?", "for", "in", ".","!", "<", ">", "+", "-", "'", "_")
```

#Creating empty lists and data frame for storing results
```{r}
#Dataframes for performance metrics
rf_df <- as.data.frame(matrix(0, ncol = 18))
names(rf_df) <- coln
#Empty lists for all predictions                         
preds_index <- list()
#Empty list for importance df
importance_df <- list()
#Empty lists for confusion matrices
rf_res <- list()
```

Looping
```{r}
#Creating list of pairs
folds <- levels(as.factor(dk_bc$.folds))
#Looping through all pairs
for(i in 1:length(folds)){
  
  #Splitting train and test data
  train.data <- dk_bc[!dk_bc$.folds==folds[i],]
  test.data <- dk_bc[dk_bc$.folds==folds[i],]
  
  #Creating DTM from test data
  freq99_t <- round(0.01*nrow(test.data)) #getting numeric value for using 99% most frequent words
  corp_t <- corpus(as.character(test.data$Transcription))
  dtm_t <- dfm(corp_t)
  dtm_t <- dfm_trim(dtm_t, min_docfreq = freq99_t, verbose = TRUE)
  dtm_t <- convert(dtm_t, to="tm")
  dtm_df_t <- as.data.frame(as.matrix(dtm_t))
  #Add binary and wordcount to dtm
  dtm_df_t$binary <- test.data$binary
  dtm_df_t$wordcount <- test.data$wordcount
  #Creating DTM from training data
  freq99 <- round(0.01*nrow(train.data))
  corp <-corpus(as.character(train.data$Transcription))
  dtm <- dfm(corp)
  dtm <- dfm_trim(dtm, min_termfreq = freq99, verbose = TRUE)
  dtm <- convert(dtm, to="tm")
  dtm_df <- as.data.frame(as.matrix(dtm))
  dtm_df <- dtm_df[,names(dtm_df) %in% names(dtm_df_t)]
  
  #Add binary and wordcount to dtm
  dtm_df$binary <- train.data$binary
  dtm_df$wordcount <- train.data$wordcount
  
  #Remove all columns random forest cant handle
  if(length(which(names(dtm_df) %in% bad_rf)) !=0){
    dtm_df <- dtm_df[,-which(names(dtm_df) %in% bad_rf)]
  }
  
  #Running Random Forest classification
  rf <- randomForest(formula = as.factor(binary)~., 
                data = dtm_df, importance=TRUE,proximity=F) #CHECK
  #Saving results
  results_df <- data.frame(binary = dtm_df_t$binary, 
                         rf_pred =predict(rf,dtm_df_t),
                         index =test.data$index) 
  
  #Making confusion matrix
  results_df$rf_pred <- factor(results_df$rf_pred, levels = c("1","0"))
  results_df$binary <- factor(results_df$binary, levels = c("1","0"))
  
  cfm_rf <- confusionMatrix(
  results_df$rf_pred,
  results_df$binary,
  positive = NULL,
  dnn = c("final", "binary"),
  prevalence = NULL,
  mode = "sens_spec"
  )
  
  #Saving performance metrics
  rf_df <- rbind(rf_df,c(cfm_rf$overall,cfm_rf$byClass))
  
  #Saving results df in list
  preds_index[[i]] <- results_df
  
  #Saving importance df
  importance_df[[i]] <- as.data.frame(round(randomForest::importance(rf), 2))
  #saving confusion matrices
  rf_res[[i]] <- cfm_rf$table
}
save(rf_df,preds_index,importance_df,rf_res, file = "bc_dk.rda")
```

######################################
"DK_BC_incl_prev"
#######################################
###Creating classifier for backchanneling in Danish
```{r}
#loading packages
library(pacman)
p_load(gtools, e1071, randomForest, RCurl, caret, tm, quanteda, tidyverse)

#loading data
dk_bc <- read.csv("df_DK.csv")
dk_bc$binary <- dk_bc$Backchannel
```

Cleaning text
```{r}
#Removing punctuation and numbers, as Random Forest can't deal with them
dk_bc$Transcription <-  gsub('[[:punct:] ]+',' ',dk_bc$Transcription)
dk_bc$Transcription <-  gsub('\\d+','tal',dk_bc$Transcription)
```

Assigning wordlists
```{r}
#Column names for result df
coln <- c("Accuracy","Kappa","AccuracyLower","AccuracyUpper","AccuracyNull","AccuracyPValue","McnemarPValue","Sensitivity","Specificity","Pos Pred Value","Neg Pred Value","Precision","Recall","F1","Prevalence","Detection Rate",
          "Detection Prevalence","Balanced Accuracy")

#words and characters the randomForest function cant handle
bad_rf <- c(")", "(", ",", "?", "for", "in", ".","!", "<", ">", "+", "-", "'", "_")
bad_rf <- c(bad_rf, paste(bad_rf, "_prev", sep=""))
```

#Creating empty lists and data frame for storing results
```{r}
#Dataframes for performance metrics
rf_df <- as.data.frame(matrix(0, ncol = 18))
names(rf_df) <- coln

#Empty lists for all predictions                         
preds_index <- list()

#Empty list for importance df
importance_df <- list()

#Empty lists for confusion matrices
rf_res <- list()
```

Looping
```{r}
#Creating list of folds
folds <- levels(as.factor(dk_bc$.folds))

#Looping through all pairs
for(i in 1:length(folds)){
  
  #Splitting train and test data
  train.data <- dk_bc[!dk_bc$.folds==folds[i],]
  test.data <- dk_bc[dk_bc$.folds==folds[i],]
  
  #Creating DTM from test data
  freq99_t <- round(0.01*nrow(test.data)) #getting numeric value for using 99% most frequent words
  corp_t <- corpus(as.character(test.data$Transcription))
  dtm_t <- dfm(corp_t)
  dtm_t <- dfm_trim(dtm_t, min_docfreq = freq99_t, verbose = TRUE)
  dtm_t <- convert(dtm_t, to="tm")
  dtm_df_t <- as.data.frame(as.matrix(dtm_t))

  #Creating DTM from training data previous sentence
  corp_prev_t <- corpus(as.character(test.data$preceeding))
  dtm_prev_t <- dfm(corp_prev_t)
  dtm_prev_t <- dfm_trim(dtm_prev_t, min_termfreq = freq99_t, verbose = TRUE)
  dtm_prev_t <- convert(dtm_prev_t, to="tm")
  dtm_df_prev_t <- as.data.frame(as.matrix(dtm_prev_t))
  names(dtm_df_prev_t) <- paste(names(dtm_df_prev_t), "_prev", sep = "")
  dtm_df_t <- cbind(dtm_df_t, dtm_df_prev_t)

  #Add binary and wordcount to dtm
  dtm_df_t$binary <- test.data$binary
  dtm_df_t$wordcount <- test.data$wordcount
  dtm_df_t$wordcount_prev <- test.data$wordcount_prev
  dtm_df_t$lex_align <- test.data$lex_align
  dtm_df_t$struc_align <- test.data$struc_align
  dtm_df_t$overlap <- test.data$overlap

  #Creating DTM from training data
  freq99 <- round(0.01*nrow(train.data))
  corp <-corpus(as.character(train.data$Transcription))
  dtm <- dfm(corp)
  dtm <- dfm_trim(dtm, min_termfreq = freq99, verbose = TRUE)
  dtm <- convert(dtm, to="tm")
  dtm_df <- as.data.frame(as.matrix(dtm))

  #Creating dtm from previous line
  corp_prev <- corpus(as.character(train.data$preceeding))
  dtm_prev <- dfm(corp_prev)
  dtm_prev <- dfm_trim(dtm_prev, min_termfreq = freq99, verbose = TRUE)
  dtm_prev <- convert(dtm_prev, to="tm")
  dtm_df_prev <- as.data.frame(as.matrix(dtm_prev))
  names(dtm_df_prev) <- paste(names(dtm_df_prev), "_prev", sep = "")
  dtm_df <- cbind(dtm_df, dtm_df_prev)
  dtm_df <- dtm_df[,names(dtm_df) %in% names(dtm_df_t)]

  #Add binary, wordcount and alignment features to dtm
  dtm_df$binary <- train.data$binary
  dtm_df$wordcount <- train.data$wordcount
  dtm_df$wordcount_prev <- train.data$wordcount_prev
  dtm_df$lex_align <- train.data$lex_align
  dtm_df$struc_align <- train.data$struc_align
  dtm_df$overlap <- train.data$overlap

  #Remove all columns random forest cant handle
  if(length(which(names(dtm_df) %in% bad_rf)) !=0){
    dtm_df <- dtm_df[,-which(names(dtm_df) %in% bad_rf)]
  }
  
  #Running Random Forest classification
  rf <- randomForest(formula = as.factor(binary)~., 
                data = dtm_df, importance=TRUE,proximity=F) #CHECK

  #Saving results
  results_df <- data.frame(binary = dtm_df_t$binary, 
                         rf_pred =predict(rf,dtm_df_t),
                         index =test.data$index) 
  
  #Making confusion matrix
  results_df$rf_pred <- factor(results_df$rf_pred, levels = c("1","0"))
  results_df$binary <- factor(results_df$binary, levels = c("1","0"))
  
  cfm_rf <- confusionMatrix(
  results_df$rf_pred,
  results_df$binary,
  positive = NULL,
  dnn = c("final", "binary"),
  prevalence = NULL,
  mode = "sens_spec"
  )
  print(paste("making random forest second bc, fold=", as.character(i), sep=" "))

  #Saving performance metrics
  rf_df <- rbind(rf_df,c(cfm_rf$overall,cfm_rf$byClass))
  
  #Saving results df in list
  preds_index[[i]] <- results_df
  
  #Saving importance df
  importance_df[[i]] <- as.data.frame(round(randomForest::importance(rf), 2))

  #saving confusion matrices
  rf_res[[i]] <- cfm_rf$table
}
print("dkprev_done")
save(rf_df,preds_index, importance_df, rf_res,file =  "bc_dk_prev.rda")

```
 
#################
title: "NO_BC"
###################
###Creating classifier for backchanneling in Danish
```{r}
#loading packages
library(pacman)
p_load(gtools, e1071, randomForest, RCurl, caret, tm, quanteda, tidyverse)

#loading data
no_bc <- read.csv("df_NO.csv")
no_bc$binary <- no_bc$Backchannel
```

Cleaning text
```{r}
#Removing punctuation and numbers, as Random Forest can't deal with them
no_bc$Transcription <-  gsub('[[:punct:] ]+',' ',no_bc$Transcription)
no_bc$Transcription <-  gsub('\\d+','tal',no_bc$Transcription)
```

Assigning wordlists
```{r}
#Column names for result df
coln <- c("Accuracy","Kappa","AccuracyLower","AccuracyUpper","AccuracyNull","AccuracyPValue","McnemarPValue","Sensitivity","Specificity","Pos Pred Value","Neg Pred Value","Precision","Recall","F1","Prevalence","Detection Rate",
          "Detection Prevalence","Balanced Accuracy")

#words and characters the randomForest function cant handle
bad_rf <- c(")", "(", ",", "?", "for", "in", ".","!", "<", ">", "+", "-", "'", "_")
```

#Creating empty lists and data frame for storing results
```{r}
#Dataframes for performance metrics
rf_df <- as.data.frame(matrix(0, ncol = 18))
names(rf_df) <- coln

#Empty lists for all predictions                         
preds_index <- list()

#Empty list for importance df
importance_df <- list()

#Empty lists for confusion matrices
rf_res <- list()
```

Looping
```{r}
#Creating list of pairs
folds <- levels(as.factor(no_bc$.folds))

#Looping through all pairs
for(i in 1:length(folds)){
  
  #Splitting train and test data
  train.data <- no_bc[!no_bc$.folds==folds[i],]
  test.data <- no_bc[no_bc$.folds==folds[i],]
  
  #Creating DTM from test data
  freq99_t <- round(0.01*nrow(test.data)) #getting numeric value for using 99% most frequent words
  corp_t <- corpus(as.character(test.data$Transcription))
  dtm_t <- dfm(corp_t)
  dtm_t <- dfm_trim(dtm_t, min_docfreq = freq99_t, verbose = TRUE)
  dtm_t <- convert(dtm_t, to="tm")
  dtm_df_t <- as.data.frame(as.matrix(dtm_t))

  #Add binary and wordcount to dtm
  dtm_df_t$binary <- test.data$binary
  dtm_df_t$wordcount <- test.data$wordcount

  #Creating DTM from training data
  freq99 <- round(0.01*nrow(train.data))
  corp <-corpus(as.character(train.data$Transcription))
  dtm <- dfm(corp)
  dtm <- dfm_trim(dtm, min_termfreq = freq99, verbose = TRUE)
  dtm <- convert(dtm, to="tm")
  dtm_df <- as.data.frame(as.matrix(dtm))
  dtm_df <- dtm_df[,names(dtm_df) %in% names(dtm_df_t)]
  
  #Add binary and wordcount to dtm
  dtm_df$binary <- train.data$binary
  dtm_df$wordcount <- train.data$wordcount
  
  #Remove all columns random forest cant handle
  if(length(which(names(dtm_df) %in% bad_rf)) !=0){
    dtm_df <- dtm_df[,-which(names(dtm_df) %in% bad_rf)]
  }
  print(paste("making random forest third bc, fold=", as.character(i), sep=" "))

  #Running Random Forest classification
  rf <- randomForest(formula = as.factor(binary)~., 
                data = dtm_df, importance=TRUE,proximity=F) #CHECK
  #Saving results
  results_df <- data.frame(binary = dtm_df_t$binary, 
                         rf_pred =predict(rf,dtm_df_t),
                         index =test.data$index) 
  
  #Making confusion matrix
  results_df$rf_pred <- factor(results_df$rf_pred, levels = c("1","0"))
  results_df$binary <- factor(results_df$binary, levels = c("1","0"))
  
  cfm_rf <- confusionMatrix(
  results_df$rf_pred,
  results_df$binary,
  positive = NULL,
  dnn = c("final", "binary"),
  prevalence = NULL,
  mode = "sens_spec"
  )
  
  #Saving performance metrics
  rf_df <- rbind(rf_df,c(cfm_rf$overall,cfm_rf$byClass))
  
  #Saving results df in list
  preds_index[[i]] <- results_df
  
  #Saving importance df
  importance_df[[i]] <- as.data.frame(round(randomForest::importance(rf), 2))

  #saving confusion matrices
  rf_res[[i]] <- cfm_rf$table
}
print("bc_no_done")
save(rf_df,preds_index, importance_df, rf_res,file =  "bc_no.rda")
```
 
############################
title: "DK_NO_incl_prev"
############################

###Creating classifier for backchanneling in Danish
```{r}
#loading packages
library(pacman)
p_load(gtools, e1071, randomForest, RCurl, caret, tm, quanteda, tidyverse)

#loading data
no_bc <- read.csv("df_NO.csv")
no_bc$binary <- no_bc$Backchannel
```

Cleaning text
```{r}
#Removing punctuation and numbers, as Random Forest can't deal with them
no_bc$Transcription <-  gsub('[[:punct:] ]+',' ',no_bc$Transcription)
no_bc$Transcription <-  gsub('\\d+','tal',no_bc$Transcription)
```

Assigning wordlists
```{r}
#Column names for result df
coln <- c("Accuracy","Kappa","AccuracyLower","AccuracyUpper","AccuracyNull","AccuracyPValue","McnemarPValue","Sensitivity","Specificity","Pos Pred Value","Neg Pred Value","Precision","Recall","F1","Prevalence","Detection Rate",
          "Detection Prevalence","Balanced Accuracy")

#words and characters the randomForest function cant handle
bad_rf <- c(")", "(", ",", "?", "for", "in", ".","!", "<", ">", "+", "-", "'", "_")
bad_rf <- c(bad_rf, paste(bad_rf, "_prev", sep=""))
```

#Creating empty lists and data frame for storing results
```{r}
#Dataframes for performance metrics
rf_df <- as.data.frame(matrix(0, ncol = 18))
names(rf_df) <- coln

#Empty lists for all predictions                         
preds_index <- list()

#Empty list for importance df
importance_df <- list()

#Empty lists for confusion matrices
rf_res <- list()
```

Looping
```{r}
#Creating list of pairs
folds <- levels(as.factor(no_bc$.folds))

#Looping through all pairs
for(i in 1:length(folds)){
  
  #Splitting train and test data
  train.data <- no_bc[!no_bc$.folds==folds[i],]
  test.data <- no_bc[no_bc$.folds==folds[i],]
  
  #Creating DTM from test data
  freq99_t <- round(0.01*nrow(test.data)) #getting numeric value for using 99% most frequent words
  corp_t <- corpus(as.character(test.data$Transcription))
  dtm_t <- dfm(corp_t)
  dtm_t <- dfm_trim(dtm_t, min_docfreq = freq99_t, verbose = TRUE)
  dtm_t <- convert(dtm_t, to="tm")
  dtm_df_t <- as.data.frame(as.matrix(dtm_t))

  #Creating DTM from training data previous sentence
  corp_prev_t <- corpus(as.character(test.data$preceeding))
  dtm_prev_t <- dfm(corp_prev_t)
  dtm_prev_t <- dfm_trim(dtm_prev_t, min_termfreq = freq99_t, verbose = TRUE)
  dtm_prev_t <- convert(dtm_prev_t, to="tm")
  dtm_df_prev_t <- as.data.frame(as.matrix(dtm_prev_t))
  names(dtm_df_prev_t) <- paste(names(dtm_df_prev_t), "_prev", sep = "")
  dtm_df_t <- cbind(dtm_df_t, dtm_df_prev_t)

  #Add binary and wordcount to dtm
  dtm_df_t$binary <- test.data$binary
  dtm_df_t$wordcount <- test.data$wordcount
  dtm_df_t$wordcount_prev <- test.data$wordcount_prev
  dtm_df_t$lex_align <- test.data$lex_align
  dtm_df_t$struc_align <- test.data$struc_align
  dtm_df_t$overlap <- test.data$overlap


  #Creating DTM from training data
  freq99 <- round(0.01*nrow(train.data))
  corp <-corpus(as.character(train.data$Transcription))
  dtm <- dfm(corp)
  dtm <- dfm_trim(dtm, min_termfreq = freq99, verbose = TRUE)
  dtm <- convert(dtm, to="tm")
  dtm_df <- as.data.frame(as.matrix(dtm))

  #Creating dtm from previous line
  corp_prev <- corpus(as.character(train.data$preceeding))
  dtm_prev <- dfm(corp_prev)
  dtm_prev <- dfm_trim(dtm_prev, min_termfreq = freq99, verbose = TRUE)
  dtm_prev <- convert(dtm_prev, to="tm")
  dtm_df_prev <- as.data.frame(as.matrix(dtm_prev))
  names(dtm_df_prev) <- paste(names(dtm_df_prev), "_prev", sep = "")
  dtm_df <- cbind(dtm_df, dtm_df_prev)
  dtm_df <- dtm_df[,names(dtm_df) %in% names(dtm_df_t)]

  #Add binary, wordcount and alignment features to dtm
  dtm_df$binary <- train.data$binary
  dtm_df$wordcount <- train.data$wordcount
  dtm_df$wordcount_prev <- train.data$wordcount_prev
  dtm_df$lex_align <- train.data$lex_align
  dtm_df$struc_align <- train.data$struc_align
  dtm_df$overlap <- train.data$overlap

  #Remove all columns random forest cant handle
  if(length(which(names(dtm_df) %in% bad_rf)) !=0){
    dtm_df <- dtm_df[,-which(names(dtm_df) %in% bad_rf)]
  }
  print(paste("making random forest fourth bc, fold=", as.character(i), sep=" "))

  #Running Random Forest classification
  rf <- randomForest(formula = as.factor(binary)~., 
                data = dtm_df, importance=TRUE,proximity=F) #CHECK

  #Saving results
  results_df <- data.frame(binary = dtm_df_t$binary, 
                         rf_pred =predict(rf,dtm_df_t),
                         index =test.data$index) 
  
  #Making confusion matrix
  results_df$rf_pred <- factor(results_df$rf_pred, levels = c("1","0"))
  results_df$binary <- factor(results_df$binary, levels = c("1","0"))
  
  cfm_rf <- confusionMatrix(
  results_df$rf_pred,
  results_df$binary,
  positive = NULL,
  dnn = c("final", "binary"),
  prevalence = NULL,
  mode = "sens_spec"
  )
  
  #Saving performance metrics
  rf_df <- rbind(rf_df,c(cfm_rf$overall,cfm_rf$byClass))
  
  #Saving results df in list
  preds_index[[i]] <- results_df
  
  #Saving importance df
  importance_df[[i]] <- as.data.frame(round(randomForest::importance(rf), 2))

  #saving confusion matrices
  rf_res[[i]] <- cfm_rf$table
}

print("bc_no_prev_done")
save(rf_df,preds_index, importance_df, rf_res,file =  "bc_no_prev.rda")

```

 
 
 
 
 
 